---
title: "When and Where do auto thefts occur?"
author: "Jessica Jia-Ni Xu, Shaunee Callender, Shi Tang, Jianhong Guo, 0103, 103-1"
subtitle: Analyze of the auto theft data from Toronto Police Service
output:
  beamer_presentation: default
  slidy_presentation: default
  ioslides_presentation: default
  widescreen: yes
---


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# echo=FALSE will stop the code chunk from appearing in the knit document
# warning=FALSE and message=FALSE will stop R messages from appearing in the knit document
library(tidyverse)
library(readr)
library(dplyr)
install.packages("kableExtra")
library(kableExtra)
auto_thefts <- read_csv("auto_thefts.csv")
neighbourhood_profiles_2016 <- read_csv("neighbourhood_profiles_2016.csv")
new_auto_thefts <- auto_thefts %>% select(Index_, occurrenceyear, occurrencemonth, occurrenceday, occurrencedayofyear, occurrencedayofweek, occurrencehour, Hood_ID, Neighbourhood, premisetype) %>% filter(occurrenceyear >= 2014) %>% mutate(occurrencedayofweek = factor(occurrencedayofweek, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")))%>% mutate(occurrencemonth = factor(occurrencemonth, levels = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")))

```

## Introduction

- **Auto thefts Data** from Toronto Police Service

- **Goal:** reduce the number of Auto Thefts

- **Focus** on the location and time

- Statistical inferences: Histograms, box plots, bar plots

- Chi-Squared tests, hypothesis tesing


## Objectives 

To answer the following questions:

- Do some **areas** have a higher incidence of auto thefts?

- Are most auto thefts carried out at a specific **time** of day?

- Do the **area** & the **time** of incident occurred independent of each other?


## Data Summary 
New Variables:

<font size="5">- **day_night** - Groups the hours of the day: Day 06:00-18:00 & Night 18:00-06:00. </font> 

<font size="5">- **daytime** - Groups the hours of the day: Morning 06:00-12:00, afternoon 12:00-18:00, evening 18:00-00:00, late night 00:00-06:00.</font>

<font size="5">- **Area** - Groups Hood ID into districts: East York, Etobicoke, North York, Old Toronto, Scarborough and York. </font>

<font size="5">- **crime_rate** - Crime rate for a specific area in 2016.</font>

## Data Summary (cont'd)
New Data Frames:

- **new_auto_thefts** - created from *auto_thefts* with variables related to the time and location, included year 2014-2018.
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4}
glimpse(new_auto_thefts)
```


## Statistical Methods
- **Hypothesis test** Relationship between auto thefts occur in day and night

- **Chi squared test** Independence between the occurrence time and location & Independence between occurrence time and premise type.

- **Bootstrap sampling** Confidence interval of proportion of auto thefts occur in day time

## Results
### Frequency versus years (2014-2018)
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4}
new_auto_thefts %>% ggplot(aes(x = occurrenceyear)) + geom_histogram(binwidth = 1, color = "black", fill = "gray") + labs(title = "Frequency versus occurrence year", x = "Occurrence year of the crime" , y = "Total frequency from 2014 to 2018")
```

<font size="5">- Auto thefts remained fairly constant from 2014 - 2017, with an increase in 2018.</font>

## Frenquency versus months
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3.5}
new_auto_thefts %>% ggplot(aes(x = occurrencemonth)) + geom_bar(color = "black", fill = "gray") + labs(title = "Frequency versus occurrence month", x = "Occurrence month of the crime" , y = "Total frequency from 2014 to 2018")

```

<font size="5">- No obvious trends between occurrence month and number of auto thefts crimes.</font>

<font size="5">- The months with the least auto theft crimes are January and February. </font>
 
## Frenquency versus day of week
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3.5}
new_auto_thefts %>% ggplot(aes(x = occurrencedayofweek)) + geom_bar(color = "black", fill = "gray") + labs(title = "Frequency versus occurrence day of week", x = "Occurrence day of week of the crime" , y = "Total frequency from 2014 to 2018")
```

<font size="5">-The number of auto theft crimes remains fairly consistent thoughout the week.</font>

##  Frequency versus occurrence hour
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3.5}
new_auto_thefts %>% ggplot(aes(x = occurrencehour)) + geom_histogram(binwidth = 1, color = "black", fill = "gray") + labs(title = "Frequency versus occurrence hour", x = "Occurrence hour of the crime" , y = "Total frequency from 2014 to 2018")
```
<font size="5">- The frequency of auto thefts begins to steadily increase after 16 o'clock.</font>

<font size="5">- There is a mode around 23 which corresponds to 11 o’clock in the evening.</font> 

## Frequency versus occurrence hour (2014 to 2018)
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4}
new_auto_thefts %>% ggplot(aes(x=occurrencehour)) + geom_histogram(binwidth = 1, color="black", fill="gray") + labs(title = "Frequency versus occurrence hour", x = "Occurrence hour of the crime" , y = "Total frequency")  + facet_wrap(~occurrenceyear)
```

<font size="5">- The occurrence hour of auto theft crimes tends to remains fairly consistent throughout the years. </font>

## Summary table (proportion of auto thefts in day and night) 
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=6}
dayandnight <- new_auto_thefts %>% select(Index_, occurrencehour) %>% mutate(daytime = ifelse(occurrencehour >= 6 & occurrencehour < 18, "Day", "Night")) 
dayandnight %>% group_by(daytime) %>% summarize(n=n(), Proportion = n/18130)

```

## Hypothesis test
<font size="5"> $H_{0}$ = 0.5 (assume that the proportion that auto thefts happen in day time is 0.5)</font>
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4}
auto<-auto_thefts %>% select(Index_, occurrenceyear, occurrencemonth, occurrenceday, occurrencedayofyear, occurrencedayofweek, occurrencehour)%>% mutate(day_night=ifelse(occurrencehour>=6 &occurrencehour<18,"day", "night"))
test_statistic<-auto%>%summarise(prop_day=sum(day_night=="day")/n(), prop_night=sum(day_night=="night")/n(),n=n())
total<-nrow(auto)
repetitions<-10000
simulated_stats<-rep(NA,repetitions)
set.seed(6862)
for (i in 1:repetitions)
{
  new_sim <-sample(c("day","night"),
                   size=total,
                   replace = TRUE)
  sim_p<-sum(new_sim=="day")/total
  simulated_stats[i]<-sim_p
}
sim<-data_frame(prop_day=simulated_stats)
sim%>% ggplot(aes(x=prop_day))+geom_histogram(colour="black",fill="grey")+labs(x="Simulated proportions of auto thefts that happened in day time if p=0.5(samples of size n=18178)")

```

<font size="5">- P_value is extremely small, very close to 0.</font>

## Bootstraps and Confidence interval
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4}
set.seed(6862)
boot_prop<-rep(NA,10000)
for (i in 1:10000)
{boot_samp<-auto %>% sample_n(size=18178, replace=TRUE)
boot_prop[i]<-as.numeric(boot_samp %>% summarise(prop_day=sum(day_night=="day")/18178))
}
boot_prop<-data_frame(prop_day=boot_prop)
ggplot(boot_prop,aes(x=prop_day))+geom_histogram(color="black",fill="grey", bins=30)+labs(title="Bootstrap distribution of proportion of auto thefts \n occurred during day time based on a sample of 18178", x="Bootstrap proportions")+ geom_vline(xintercept = quantile(boot_prop$prop_day,c(0.025)),color="red"
)+geom_vline(xintercept = quantile(boot_prop$prop_day,c(0.975)),color="blue"
)
quantile(boot_prop$prop_day,c(0.025,0.975))

```

## Location versus auto theft crimes
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4}

table1 <- new_auto_thefts %>% group_by(Hood_ID) %>% summarize(number_of_crimes=n(), proportion=(number_of_crimes/18561)) %>% arrange(desc(number_of_crimes))
head(table1, 10)

```

<font size="5">- The highest number of crimes occured in Hood_ID 1 and 14.</font>

## District and auto theft crimes.
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4}

new_auto_thefts <- new_auto_thefts %>% mutate(Area = ifelse(Hood_ID <= 20, "Etobicoke", ifelse(21<=Hood_ID & Hood_ID<=53, "North York", ifelse(54<=Hood_ID & Hood_ID<=61,"East York",ifelse(62<=Hood_ID & Hood_ID<=105,"Old Toronto",ifelse(106<=Hood_ID & Hood_ID<=115,"York","Scarborough"))))))

table2 <- new_auto_thefts %>% group_by(Area) %>% summarize(n=n(), proportion=(n/18561)) %>% arrange(desc(n))
new_auto_thefts %>% ggplot(aes(x = Area)) + geom_bar()+labs(title="District vs Auto Theft Frequency", x="District", y= "Frequency")

```

<font size="5">- North York had the most auto theft crimes, while East York had the least.</font>

## District and Crime Rate
```{r, echo=FALSE, message=FALSE, warning=FALSE, }
table4 <- new_auto_thefts %>% filter(occurrenceyear == 2016) %>% group_by(Hood_ID) %>% summarise(n=n())
new_data <- inner_join(x = table4, y = neighbourhood_profiles_2016, by = "Hood_ID")
data_pop <- new_data %>% mutate(Area = ifelse(Hood_ID <= 20, "Etobicoke", ifelse(21<=Hood_ID & Hood_ID<=53, "North York", ifelse(54<=Hood_ID & Hood_ID<=61,"East York",ifelse(62<=Hood_ID & Hood_ID<=105,"Old Toronto",ifelse(106<=Hood_ID & Hood_ID<=115,"York","Scarborough")))))) 

data_pop <-data_pop %>% select(Hood_ID, n, pop_2016, Area)
table3 <- data_pop %>% group_by(Area) %>% summarise(n_2016=sum(n),population_2016=sum(pop_2016))
table3 <- table3 %>% mutate(crime_rate = round(((n_2016/population_2016)*100000),2))%>% arrange(desc(crime_rate))

kable(table3, caption = "Crime Rate for 2016") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  kableExtra::kable_styling(latex_options = "hold_position")

```

## Chi-squared test for time of day and district
```{r, echo=FALSE, message=FALSE, warning=FALSE}
new_auto_thefts <- new_auto_thefts %>% mutate(daytime = ifelse(occurrencehour < 6, "Late night", ifelse(occurrencehour < 12, "Morning", ifelse(occurrencehour < 18, "Afternoon", "Night"))))

areatimetable <- table(new_auto_thefts$Area, new_auto_thefts$daytime)
areatimetable

chisq.test(areatimetable)
```

## Chi-sqaured test for time of day and premise type

```{r, echo=FALSE, message=FALSE, warning=FALSE}

premisetimetable <- table(new_auto_thefts$premisetype, new_auto_thefts$daytime)
premisetimetable

chisq.test(premisetimetable)

```

## Distribution across location

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 5, fig.width = 4}
new_auto_thefts <- new_auto_thefts %>% mutate(Area = ifelse(Hood_ID <= 20, "Etobicoke", ifelse(21<=Hood_ID & Hood_ID<=53, "North York", ifelse(54<=Hood_ID & Hood_ID<=61,"East York",ifelse(62<=Hood_ID & Hood_ID<=105,"Old Toronto",ifelse(106<=Hood_ID & Hood_ID<=115,"York","Scarborough"))))))

new_auto_thefts %>% ggplot(aes(x = Area, y = occurrencehour)) + geom_boxplot() + coord_flip() + labs(title = "Occurrence hour vs District", y = "Occurrence hour of auto theft", x = "District")

new_auto_thefts %>% ggplot(aes(x = premisetype, y = occurrencehour)) + geom_boxplot() + coord_flip() + labs(title = "Occurrence hour vs Premise", y = "Occurrence hour of auto theft", x = "Premise")
```

## Conclusion

- Certain areas have a higher incidence of auto thefts which is shown by the varying crime rates.

- Most auto thefts tend to be carried out at night, more specifically between evening and late night. This was confirmed by bootstrap testing and hypothesis testing.

- There is strong evidence to suggest the time of day of the incident and the location of the incident are interdependent.


## Recommendations

- The Toronto Police Service should send more patrols to areas with a higher crime rate.

- Officers should be put on patrol at night because that is when more auto thefts tend to be carried out.


## Limitations

- As the seasons change the sun sets at different times and this may have an impact on auto thefts.

- The location data had to be generalised and this resulted in large variations in the size of areas, which affected our analysis.

- The type of car stolen was not given. Newer cars have are built differently and could possibly be easier to break into. 


